---
title: "01-exploratory-analyses"
author: "Justin Ephraim Isip"
date: "13/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(tidyverse)
library(reshape2)
library(hrbrthemes)
library(ggpubr)
library(patchwork)
```


# The 9 data arrangements are as follows:
# 1. max_bf_overall:  maximum bite force for each species (overall = males and females combined)
# 2. max_bf_males: maximum bite force for each species (males only)
# 3. max_bf_females: maximum bite force for each species (females only)
# 4. avg_bf_withsd_overall:  average bite force + standard deviation for each species (overall = males and females combined)
# 5. avg_bf_withsd_males: average bite force + standard deviation for each species (males only)
# 6. avg_bf_withsd_females: average bite force + standard deviation for each species (females only)
# 7. avg_bf_withse_overall:  average bite force + standard error for each species (overall = males and females combined)
# 8. avg_bf_withse_males: average bite force + standard error for each species (males only)
# 9. Avg_bf_withse_females: average bite force + standard error for each species (females only)


```{r}
# Read in the 9 datasets
# For female datasets, Define sex as a factor as otherwise treats F as false

max_bf_overall <- read_csv("../data/max_bf_overall.csv")
max_bf_males <- read_csv("../data/max_bf_males.csv")
max_bf_females <- read_csv("../data/max_bf_females.csv", col_types = cols(Sex = col_factor()))
avg_bf_plus_se_overall <- read_csv("../data/avg_bf_plus_se_overall.csv")
avg_bf_plus_se_males <- read_csv("../data/avg_bf_plus_se_males.csv")
avg_bf_plus_se_females <- read_csv("../data/avg_bf_plus_se_females.csv", col_types =  cols(Sex = col_factor()))
avg_bf_plus_sd_overall <- read_csv("../data/avg_bf_plus_sd_overall.csv")
avg_bf_plus_sd_males <- read_csv("../data/avg_bf_plus_sd_males.csv")
avg_bf_plus_sd_females <- read_csv("../data/avg_bf_plus_sd_females.csv", col_types =  cols(Sex = col_factor()))

```

First, let's combine males with females, so we can compare between sexes when needed
```{r}
max_bf_mf_combined <- bind_rows(max_bf_males, max_bf_females)
avg_bf_plus_se_mf_combined <- bind_rows(avg_bf_plus_se_males, avg_bf_plus_se_females)
avg_bf_plus_sd_mf_combined <- bind_rows(avg_bf_plus_sd_males, avg_bf_plus_sd_females)
```

What’s the distribution of bite forces overall for each dataset? (histograms)
```{r}
ggplot(max_bf_overall, aes(x=log(max_bf), alpha= 0.5)) + geom_histogram(binwidth=0.5) +  
  ggtitle("Distribution of max bf overall (M + F)")
ggplot(max_bf_males, aes(x=log(max_bf), alpha= 0.5)) + geom_histogram(binwidth=0.5) +  
  ggtitle("Distribution of max bf (M only)")
ggplot(max_bf_females, aes(x=log(max_bf), alpha= 0.5)) + geom_histogram(binwidth=0.5) + 
  ggtitle("Distribution of max bf (F only)")
ggplot(avg_bf_plus_se_overall, aes(x=log(bf_plus_se), alpha= 0.5)) + geom_histogram(binwidth=0.5) +  
  ggtitle("Distribution of BF + SE overall (M + F)")
ggplot(avg_bf_plus_se_males, aes(x=log(bf_plus_se), alpha= 0.5)) + geom_histogram(binwidth=0.5)+  
  ggtitle("Distribution of BF + SE (M only)")
ggplot(avg_bf_plus_se_females, aes(x=log(bf_plus_se), alpha= 0.5)) + geom_histogram(binwidth=0.5) +  
  ggtitle("Distribution of BF + SE (F only)")
ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd), alpha= 0.5)) + geom_histogram(binwidth=0.5) +  
  ggtitle("Distribution of BF + SD overall (M+F)")
ggplot(avg_bf_plus_sd_males, aes(x=log(bf_plus_sd), alpha= 0.5)) + geom_histogram(binwidth=0.5) +  
  ggtitle("Distribution of BF + SD (M only)")
ggplot(avg_bf_plus_sd_females, aes(x=log(bf_plus_sd), alpha= 0.5)) + geom_histogram(binwidth=0.5) +  
  ggtitle("Distribution of BF + SD (F only)")

```

What’s the distribution of bite forces overall for each dataset? (density plots)
```{r}
  ggplot(max_bf_overall, aes(x=log(max_bf))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of max bf overall (M + F)")

  ggplot(max_bf_males, aes(x=log(max_bf))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of max bf (males)")
  
    ggplot(max_bf_females, aes(x=log(max_bf))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of max bf (females)")
    
      ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of BF + SD (overall)")
      
        ggplot(avg_bf_plus_sd_males, aes(x=log(bf_plus_sd))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of BF + SD (males)")
        
    ggplot(avg_bf_plus_sd_females, aes(x=log(bf_plus_sd))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of BF + SD (females)")
    
    ggplot(avg_bf_plus_se_overall, aes(x=log(bf_plus_se))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of BF + SE (overall)")
        
    ggplot(avg_bf_plus_se_males, aes(x=log(bf_plus_se))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of BF + SD (males)")
        
    ggplot(avg_bf_plus_se_females, aes(x=log(bf_plus_se))) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
    ggtitle("Distribution of BF + SD (females)")
```

What's the distribution of bite forces between males and females for MAX BF, BF + SD and BF + SE? (histograms)
```{r}

ggplot(max_bf_mf_combined, aes(x = log(max_bf), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.5, alpha = 0.6) +
  ggtitle("Distribution of MAX BF of males and females")

ggplot(avg_bf_plus_se_mf_combined, aes(x = log(bf_plus_se), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.5, alpha = 0.6) +
  ggtitle("Distribution of BF + SE of males and females")

ggplot(avg_bf_plus_sd_mf_combined, aes(x = log(bf_plus_sd), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.5, alpha = 0.6) +
  ggtitle("Distribution of BF + SD of males and females")

# Males have slightly higher bite forces on average across all 3 data arrangements
```

What's the distribution of bite forces between males and females for MAX BF, BF + SD and BF + SE? (density plots)
```{r}
ggplot(data=max_bf_mf_combined, aes(x=log(max_bf), group=Sex, fill=Sex)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of MAX BF of males and females")

ggplot(data=avg_bf_plus_sd_mf_combined, aes(x=log(bf_plus_sd), group=Sex, fill=Sex)) +
    geom_density(adjust=1.5, alpha=.4)+
  ggtitle("Distribution of BF + SD of males and females")


ggplot(data=avg_bf_plus_se_mf_combined, aes(x=log(bf_plus_se), group=Sex, fill=Sex)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of BF + SE of males and females")
# This should be included in supp info
```

What's the distribution of SVL between males and females for MAX BF, BF + SD and BF + SE? (density plots)
```{r}
ggplot(data=max_bf_mf_combined, aes(x=log(max_svl), group=Sex, fill=Sex)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of MAX SVL of males and females")

ggplot(data=avg_bf_plus_sd_mf_combined, aes(x=log(svl_plus_sd), group=Sex, fill=Sex)) +
    geom_density(adjust=1.5, alpha=.4)+
  ggtitle("Distribution of SVL + SD of males and females")


ggplot(data=avg_bf_plus_se_mf_combined, aes(x=log(svl_plus_se), group=Sex, fill=Sex)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of SVL + SE of males and females")

```

How are bite forces distributed across our datasets based on diet?
```{r}

max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_bf), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of max_bf based on diet")

avg_bf_plus_sd_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(bf_plus_sd), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of bf + sd based on diet")

avg_bf_plus_se_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(bf_plus_se), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of bf + se based on diet")

# Density plots show that herbivores have higher bite forces
```


How are bite forces distributed across our datasets based on Main Biogeographic Realm?
```{r}

max_bf_overall %>%
  filter(!is.na(MainBiogeographicRealm)) %>%
  ggplot(aes(x=log(max_bf), group=MainBiogeographicRealm, fill=MainBiogeographicRealm, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of max bf based on MBR")


avg_bf_plus_sd_overall %>%
  filter(!is.na(MainBiogeographicRealm)) %>%
  ggplot(aes(x=log(bf_plus_sd), group=MainBiogeographicRealm, fill=MainBiogeographicRealm, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of bf + sd based on MBR")

avg_bf_plus_se_overall %>%
  filter(!is.na(MainBiogeographicRealm)) %>%
  ggplot(aes(x=log(bf_plus_se), group=MainBiogeographicRealm, fill=MainBiogeographicRealm, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of bf + se based on MBR")
```

How are bite forces distributed across our datasets based on lifestyle?
```{r}

max_bf_overall %>%
  filter(!is.na(Lifestyle)) %>%
  ggplot(aes(x=log(max_bf), group=Lifestyle, fill=Lifestyle, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of max bf based on Lifestyle")


avg_bf_plus_sd_overall %>%
  filter(!is.na(Lifestyle)) %>%
  ggplot(aes(x=log(bf_plus_sd), group=Lifestyle, fill=Lifestyle, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of bf + sd based on Lifestyle")

avg_bf_plus_se_overall %>%
  filter(!is.na(Lifestyle)) %>%
  ggplot(aes(x=log(bf_plus_se), group=Lifestyle, fill=Lifestyle, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) +
  ggtitle("Distribution of bf + se based on Lifestyle")

```

What's the distribution of SVL across males and females for the 3 data arrangements? (histogram)
```{r}

ggplot(max_bf_mf_combined, aes(x = log(max_svl), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.1, alpha = 0.6) 

ggplot(avg_bf_plus_se_mf_combined, aes(x = log(svl_plus_se), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.1, alpha = 0.6) 

ggplot(avg_bf_plus_sd_mf_combined, aes(x = log(svl_plus_sd), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.1, alpha = 0.6) 

```

What's the relationship between BF and all morphometric variables based on sex? (max bf combined)
```{r}
ggplot(max_bf_mf_combined, aes(x=log(max_bf), y =log(max_svl), shape = Sex, colour = Sex)) + 
  geom_point() 

ggplot(max_bf_mf_combined, aes(x=log(max_bf), y =log(max_bm), shape = Sex, colour = Sex)) + 
  geom_point() 

ggplot(max_bf_mf_combined, aes(x=log(max_bf), y =log(max_hl), shape = Sex, colour = Sex)) + 
  geom_point() 

ggplot(max_bf_mf_combined, aes(x=log(max_bf), y =log(max_hh), shape = Sex, colour = Sex)) + 
  geom_point() 

ggplot(max_bf_mf_combined, aes(x=log(max_bf), y =log(max_hw), shape = Sex, colour = Sex)) + 
  geom_point() 

ggplot(max_bf_mf_combined, aes(x=log(max_bf), y =log(max_ljl), shape = Sex, colour = Sex)) + 
  geom_point() 

#There is a positive correlation between BF and all morphometric variables, however there are no clear patterns between sexes
```

What's the relationship between BF and all morphometric variables based on Family?	(max bf overall)
```{r}
ggplot(max_bf_overall, aes(x=log(max_bf), y =log(max_svl), shape = Family, colour = Family)) + 
  geom_point(alpha = 0.5) +
    scale_shape_manual(values=seq(0,20))

ggplot(max_bf_overall, aes(x=log(max_bf), y =log(max_bm), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(max_bf_overall, aes(x=log(max_bf), y =log(max_hl), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(max_bf_overall, aes(x=log(max_bf), y =log(max_hh), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(max_bf_overall, aes(x=log(max_bf), y =log(max_hw), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(max_bf_overall, aes(x=log(max_bf), y =log(max_ljl), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))
```


What's the relationship between BF and all morphometric variables based on Family?	(avg_bf_plus_sd_overall)
```{r}
ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd), y =log(svl_plus_sd), shape = Family, colour = Family)) + 
  geom_point(alpha = 0.5) +
    scale_shape_manual(values=seq(0,20))

ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd), y =log(bm_plus_sd), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd), y =log(hl_plus_sd), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd), y =log(hh_plus_sd), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd), y =log(hw_plus_sd), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggplot(avg_bf_plus_sd_overall, aes(x=log(bf_plus_sd), y =log(ljl_plus_sd), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))
```

Patchwork grid of distribution of all variables (max_bf_overall dataset)
```{r}
#library(patchwork)
p1 <-  ggplot(max_bf_overall, aes(x=log(max_bf))) +
    geom_density(color="blue", fill="blue", alpha=0.8) +
    ggtitle("Distribution of max bf")

p2 <-  ggplot(max_bf_overall, aes(x=log(max_svl))) +
    geom_density(colour = "red", fill = "red", alpha=0.8) +
    ggtitle("Distribution of max svl")

p3 <-  ggplot(max_bf_overall, aes(x=log(max_bm))) +
    geom_density(colour = "green", fill = "green",  alpha=0.8) +
    ggtitle("Distribution of max bm")

p4 <-  ggplot(max_bf_overall, aes(x=log(max_hl))) +
    geom_density(colour = "yellow", fill = "yellow",  alpha=0.8) +
    ggtitle("Distribution of max hl")

p5 <-  ggplot(max_bf_overall, aes(x=log(max_hh))) +
    geom_density(colour = "purple", fill = "purple",  alpha=0.8) +
    ggtitle("Distribution of max hh")

p6 <-  ggplot(max_bf_overall, aes(x=log(max_hw))) +
    geom_density(colour = "brown", fill = "brown",  alpha=0.8) +
    ggtitle("Distribution of max hw")

p7 <-  ggplot(max_bf_overall, aes(x=log(max_ljl))) +
    geom_density(colour = "orange", fill = "orange",  alpha=0.8) +
    ggtitle("Distribution of max ljl")



p1 + p2 + p3 + p4 + p5 + p6 + p7 + plot_layout(nrow = 3, byrow = F)
```
	
	
How are all of the variables distributed based on diet?
```{r}
p8 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_bf), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) 
  

p9 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_svl), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) 
 

p10 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_bm), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) 


p11 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_hl), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) 
 

p12 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_hw), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) 


p13 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_hh), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) 

p14 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_ljl), group=MieriDiet, fill=MieriDiet, na.rm = T)) +
    geom_density(adjust=1.5, alpha=.4) 



ggarrange(p8 , p9 , p10 , p11 ,p12 ,p13 , p14, ncol = 3, nrow = 3, common.legend = TRUE, legend ="bottom")  +  plot_annotation(title = "Distribution of all the variables based on diet")


```


How are all the variables distributed based on sex?
```{r}

p15 <- max_bf_mf_combined %>%
  ggplot(aes(x=log(max_bf), group=Sex, fill=Sex)) +
    geom_density( alpha= 0.3) +
  theme_bw() +
  xlab("log(BF) (N)")+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)) +
  scale_fill_discrete(name = "Sex", labels = c("Female", "Male"))


p16<- max_bf_mf_combined %>%
  ggplot(aes(x=log(max_svl), group=Sex, fill=Sex)) +
    geom_density(alpha=0.3) +
  theme_bw() +
  xlab("log(SVL) (mm)") +
  ylim(0,1.5) +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))
 

p17 <- max_bf_mf_combined %>%
  ggplot(aes(x=log(max_bm), group=Sex, fill=Sex)) +
    geom_density(alpha=0.3) +
  theme_bw() +
  xlab("log(BM) (g)") +
  ylim(0, 0.3) +
  xlim(0,6)+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))


p18 <- max_bf_mf_combined %>%
  ggplot(aes(x=log(max_hl), group=Sex, fill=Sex)) +
  geom_density(alpha=0.3) +
  theme_bw() +
  xlab("log(HL) (mm)") +
  ylim(0,1.5)+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

 

p19 <- max_bf_mf_combined %>%
  ggplot(aes(x=log(max_hw), group=Sex, fill=Sex)) +
    geom_density(alpha=0.3) +
  theme_bw() +
  ylim(0,1.5) +
  xlab("log(HW) (mm)") +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))


p20 <- max_bf_mf_combined %>%
  ggplot(aes(x=log(max_hh), group=Sex, fill=Sex)) +
    geom_density(alpha=0.3) +
  theme_bw() +
  ylim(0,1.5) +
  xlab("log(HH) (mm)") +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))
 

p21 <- max_bf_mf_combined %>%
  ggplot(aes(x=log(max_ljl), group=Sex, fill=Sex)) +
    geom_density(alpha=0.3) +
  theme_bw() +
  ylim(0,1.5) +
  xlab("log(LJL) (mm)") +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))



ggarrange(p15,p16,p17,p18,p19,p20,p21, ncol = 2, nrow = 4, common.legend = TRUE, legend ="bottom")

ggsave("../figures/male-female-density-plots.png", dpi = 300)
```


How are all the variables distributed based on families?
```{r}

p22 <- max_bf_overall %>%
  ggplot(aes(x=log(max_bf), group=Family, fill=Family)) +
    geom_density(adjust=1.5, alpha=.4) +
facet_wrap(~Family)
  

p23<- max_bf_overall %>%
  ggplot(aes(x=log(max_svl), group=Family, fill=Family)) +
    geom_density(adjust=1.5, alpha=.4)  +
facet_wrap(~Family)
 

p24 <- max_bf_overall %>%
  ggplot(aes(x=log(max_bm), group=Family, fill=Family)) +
    geom_density(adjust=1.5, alpha=.4) 


p25 <- max_bf_overall %>%
  ggplot(aes(x=log(max_hl), group=Family, fill=Family)) +
    geom_density(adjust=1.5, alpha=.4) 
 

p26 <- max_bf_overall %>%
  ggplot(aes(x=log(max_hw), group=Family, fill=Family)) +
    geom_density(adjust=1.5, alpha=.4) 


p27 <- max_bf_overall %>%
  ggplot(aes(x=log(max_hh), group=Family, fill=Family)) +
    geom_density(adjust=1.5, alpha=.4) 

p28 <- max_bf_overall %>%
  ggplot(aes(x=log(max_ljl), group=Family, fill=Family)) +
    geom_density(adjust=1.5, alpha=.4) 



ggarrange(p22, p23 , p24 , p25, ncol = 2, nrow = 2, common.legend = TRUE, legend ="bottom")  +  plot_annotation(title = "Distribution of all the variables based on Family")


ggarrange(p26 ,p27 ,p28, ncol = 2, nrow = 2, common.legend = TRUE, legend ="bottom")  +  plot_annotation(title = "Distribution of all the variables based on Family")


```

How are all the variables distributed based on lifestyle?
```{r}

life1 <- max_bf_overall %>%
  ggplot(aes(x=log(max_bf), group=Lifestyle, fill=Lifestyle)) +
    geom_density(adjust=1.5, alpha=.4) +
facet_wrap(~Lifestyle)
  

life2 <- max_bf_overall %>%
  ggplot(aes(x=log(max_svl), group=Lifestyle, fill=Lifestyle)) +
    geom_density(adjust=1.5, alpha=.4)  +
facet_wrap(~Lifestyle)
 

life3 <- max_bf_overall %>%
  ggplot(aes(x=log(max_bm), group=Lifestyle, fill=Lifestyle)) +
    geom_density(adjust=1.5, alpha=.4)  +
facet_wrap(~Lifestyle)


life4 <- max_bf_overall %>%
  ggplot(aes(x=log(max_hl), group=Lifestyle, fill=Lifestyle)) +
    geom_density(adjust=1.5, alpha=.4)  +
facet_wrap(~Lifestyle)
 

life5 <- max_bf_overall %>%
  ggplot(aes(x=log(max_hw), group=Lifestyle, fill=Lifestyle)) +
    geom_density(adjust=1.5, alpha=.4)  +
facet_wrap(~Lifestyle)


life6 <- max_bf_overall %>%
  ggplot(aes(x=log(max_hh), group=Lifestyle, fill=Lifestyle)) +
    geom_density(adjust=1.5, alpha=.4)  +
facet_wrap(~Lifestyle)

life7 <- max_bf_overall %>%
  ggplot(aes(x=log(max_ljl), group=Lifestyle, fill=Lifestyle)) +
    geom_density(adjust=1.5, alpha=.4)  +
facet_wrap(~Lifestyle)



life1
life2
life3
life4
life5
life6
life7

```


What's the relationship between BF and all morphometric variables based on diet?	(max bf)
```{r}
  
p29 <- max_bf_overall %>% 
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_svl), y =log(max_bf), shape = MieriDiet, colour = MieriDiet)) + 
  geom_point() 

p30 <-  max_bf_overall %>% 
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_bm), y =log(max_bf), shape = MieriDiet, colour = MieriDiet)) + 
  geom_point() 

p31 <- max_bf_overall %>% 
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_hl), y =log(max_bf), shape = MieriDiet, colour = MieriDiet)) + 
  geom_point() 

p32 <- max_bf_overall %>% 
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_hh), y =log(max_bf), shape = MieriDiet, colour = MieriDiet)) + 
  geom_point() 

p33 <-  max_bf_overall %>% 
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_hw), y =log(max_bf), shape = MieriDiet, colour = MieriDiet)) + 
  geom_point() 

p34 <- max_bf_overall %>% 
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=log(max_ljl), y =log(max_bf), shape = MieriDiet, colour = MieriDiet)) + 
  geom_point() 

ggarrange( p29 , p30 , p31 ,p32 ,p33 ,p34,  ncol = 3, nrow = 2, common.legend = TRUE, legend ="bottom")  +  plot_annotation(title = "Relationship between BF and all morphological variables")



```

## What's the relationship between BF and all morphometric variables based on Family?	(max bf)
	
```{r}
  
p41 <- max_bf_overall %>% 
  ggplot(aes(x=log(max_svl), y =log(max_bf), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

p42 <-  max_bf_overall %>% 
  ggplot(aes(x=log(max_bm), y =log(max_bf), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

p43 <- max_bf_overall %>% 
  ggplot(aes(x=log(max_hl), y =log(max_bf), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

p44 <- max_bf_overall %>% 
  ggplot(aes(x=log(max_hh), y =log(max_bf), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

p45 <-  max_bf_overall %>% 
  ggplot(aes(x=log(max_hw), y =log(max_bf), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

p46 <- max_bf_overall %>% 
  ggplot(aes(x=log(max_ljl), y =log(max_bf), shape = Family, colour = Family)) + 
  geom_point() +
    scale_shape_manual(values=seq(0,20))

ggarrange( p41 , p42, ncol = 2, common.legend = TRUE, legend ="bottom")  +  plot_annotation(title = "Relationship between BF and all morphological variables")

ggarrange( p43 , p44, ncol = 2, common.legend = TRUE, legend ="bottom")  +  plot_annotation(title = "Relationship between BF and all morphological variables")

ggarrange(p45 ,p46,   ncol = 2, common.legend = TRUE, legend ="bottom")  +  plot_annotation(title = "Relationship between BF and all morphological variables")

```

## How is bite force distributed based on family?
```{r}

p35 <-ggplot(max_bf_overall,aes(x=Family,y=log(max_bf),color=Family))+
  geom_boxplot(show.legend=F,fill="white")+
  geom_jitter(show.legend=F,alpha=0.4) +
  xlab("Family") + ylab("log(max_bf)") +
  ggtitle("Distribution of max bf based on Family") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.spacing = unit(0, "lines")
  )

p36 <- ggplot(max_bf_overall,aes(x=Family,y=log(max_svl),color=Family))+
  geom_boxplot(show.legend=F,fill="white")+
  geom_jitter(show.legend=F,alpha=0.4) +
  xlab("Family") + ylab("log(max_svl)") +
  ggtitle("Distribution of max svl based on Family") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.spacing = unit(0, "lines")
  )



p37 <- max_bf_overall %>%
  ggplot(aes(x=Lifestyle,y=log(max_bf),color=Lifestyle))+
  geom_boxplot(show.legend=F,fill="white")+
  geom_jitter(show.legend=F,alpha=0.2) +
  ggtitle("Distribution of max bf based on lifestyle")+
  facet_grid(.~Lifestyle,scales="free",space="free_x") +
  theme(axis.title.x = element_blank())

p38 <- max_bf_overall %>%
  ggplot(aes(x=MainBiogeographicRealm,y=log(max_bf),color=MainBiogeographicRealm))+
  geom_boxplot(show.legend=F,fill="white")+
  geom_jitter(show.legend=F,alpha=0.2) +
  ggtitle("Distribution of max bf based on MBR")+
  facet_grid(.~MainBiogeographicRealm, scales="free",space="free_x") +
   theme(axis.title.x = element_blank())

p39 <- avg_bf_plus_sd_mf_combined %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=MieriDiet,y=log(bf_plus_sd),color=Sex))+
  geom_boxplot(show.legend=T,fill="white")+
  geom_jitter(show.legend=F,alpha=0.2) +
  ggtitle("Distribution of max bf between sexes based on diet")+
  facet_grid(.~MainBiogeographicRealm, scales="free",space="free_x") +
   theme(axis.title.x = element_blank())


p40 <- max_bf_overall %>%
  filter(!is.na(MieriDiet)) %>%
  ggplot(aes(x=MieriDiet,y=log(max_bf),color=MieriDiet))+
  geom_boxplot(show.legend=T,fill="white")+
  geom_jitter(show.legend=F,alpha=0.2) +
  ggtitle("Distribution of max bf based on diet")+
  facet_grid(.~MieriDiet, scales="free",space="free_x") +
   theme(axis.title.x = element_blank())


p35
p36
p37
p38
p39
p40
```


	