---
title: "01-exploratory-analyses"
author: "Justin Ephraim Isip"
date: "13/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

library(tidyverse)

# The 9 data arrangements are as follows:
# 1. A1:  maximum bite force for each species (overall, males and females)
# 2. A1M: maximum bite force for each species (males only)
# 3. A1F: maximum bite force for each species (females only)
# 4. A2:  average bite force + standard deviation for each species (overall, males and females)
# 5. A2M: average bite force + standard deviation for each species (males only)
# 6. A2F: average bite force + standard deviation for each species (females only)
# 7. A3:  average bite force + standard error for each species (overall, males and females)
# 8. A3M: average bite force + standard error for each species (males only)
# 9. A3F: average bite force + standard error for each species (females only)

```{r}
# Read in the 9 datasets
# Define Sex as a factor as otherwise treats F as false
A1  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A1.csv", col_types = cols( Sex = col_factor())) 
A1M  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A1M.csv", col_types = cols( Sex = col_factor())) 
A1F  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A1F.csv", col_types = cols( Sex = col_factor())) 
A2  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A2.csv", col_types = cols( Sex = col_factor())) 
A2M  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A2M.csv", col_types = cols( Sex = col_factor())) 
A2F  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A2F.csv", col_types = cols( Sex = col_factor())) 
A3  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A3.csv", col_types = cols( Sex = col_factor())) 
A3M  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A3M.csv", col_types = cols( Sex = col_factor())) 
A3F  <- read_csv("C:/Users/Justin Isip/Documents/Biteforce-lepidosaurs/data/A3F.csv", col_types = cols( Sex = col_factor())) 
```

## What’s the distribution of bite forces for each dataset?

```{r}
ggplot(A1, aes(x=log(MAX.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A1M, aes(x=log(MAX.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A1F, aes(x=log(MAX.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A2, aes(x=log(SD.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A2M, aes(x=log(SD.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A2F, aes(x=log(SD.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A3, aes(x=log(SE.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A3M, aes(x=log(SE.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)
ggplot(A3F, aes(x=log(SE.BF), alpha= 0.5)) + geom_histogram(binwidth=0.5)

```


## What's the distribution of bite forces across males and females for MAX.BF, SD.BF and SE.BF?

```{r}
# First bind the male df with female df
A1MF<- bind_rows(A1F, A1M)
A2MF<- bind_rows(A2F, A2M)
A3MF<- bind_rows(A3F, A3M)

ggplot(A1MF, aes(x = log(MAX.BF), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.5, alpha = 0.6) 

ggplot(A2MF, aes(x = log(SD.BF), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.5, alpha = 0.6) 

ggplot(A3MF, aes(x = log(SE.BF), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.5, alpha = 0.6) 

```

## What's the distribution of SVL across males and females for the 3 data arrangements?

```{r}
ggplot(A1MF, aes(x = log(MAX.SVL), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.1, alpha = 0.6) 

ggplot(A2MF, aes(x = log(SD.SVL), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.1, alpha = 0.6) 

ggplot(A3MF, aes(x = log(SE.SVL), fill = Sex)) +
  geom_histogram(position = "identity", binwidth = 0.1, alpha = 0.6) 

```


## What's the relationship between BF and all morphometric variables between males and females?

```{r}
ggplot(A1MF, aes(x=log(MAX.BF), y =log(MAX.SVL), shape = Sex, colour = Sex)) + 
  geom_point() 
  
ggplot(A1MF, aes(x=log(MAX.BF), y =log(MAX.BM), shape = Sex, colour = Sex)) + 
geom_point() 

ggplot(A1MF, aes(x=log(MAX.BF), y =log(MAX.HL), shape = Sex, colour = Sex)) + 
geom_point() 

ggplot(A1MF, aes(x=log(MAX.BF), y =log(MAX.HW), shape = Sex, colour = Sex)) + 
geom_point() 

ggplot(A1MF, aes(x=log(MAX.BF), y =log(MAX.HH),shape = Sex, colour = Sex)) + 
geom_point() 

ggplot(A1MF, aes(x=log(MAX.BF), y =log(MAX.LJL),shape = Sex, colour = Sex)) + 
geom_point() 
```
	
	