---
title: "combine-data-with-phylogeny"
author: "Justin Ephraim Isip"
date: "23/09/2020"
output: html_document
---


This script combines my data and phylogeny to conduct PGLS analysis

## Load libraries
```{r, message = FALSE, warning = FALSE}
library(ape)
library(geiger)
library(picante)
library(caper)
library(tidyverse)
library(rlist)
library(phytools)
library(ggpubr)
library(ggfortify)
library(patchwork)
library(scales)
```

## Read in my data and phylogeny
```{r, message = FALSE, warning = FALSE}
max_bf_overall <- read.csv("../data/max_bf_overall.csv", stringsAsFactors = T)
lepidosaurtree <- read.tree("../WrightTree.nex")
```


## Let's have a quick look at the tree
```{r, message = FALSE, warning = FALSE}
lepidosaurtree
str(lepidosaurtree)
```

## My tree has 4162 species and 4161 internal nodes. 
## Lets check whether my tree is dichomotomous?
```{r}
is.binary.tree(lepidosaurtree)
```

## First, lets replace spaces with underscores (_) so they match up with species names in the tree
```{r}
max_bf_overall$BinomialReptileDatabase <- gsub(" ", "_", max_bf_overall$BinomialReptileDatabase)
```

## Now, lets check mismatches between species in my data and phylogeny
```{r}
check <- name.check(phy = lepidosaurtree, data = max_bf_overall, data.names = max_bf_overall$BinomialReptileDatabase)
```

## Here I relace more recent species names that are in my dataset to match older species names that are in my phylogeny
```{r}

lepidosaurtree$tip.label <- gsub("Typhlosaurus_lineatus", "Acontias_kgalagadi", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Glaphyromorphus_isolepis", "Eremiascincus_isolepis", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Cordylus_cataphractus", "Ouroborus_cataphractus", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Cordylus_warreni", "Smaug_mossambicus", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Uromastyx_acanthinura", "Uromastix_acanthinura", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Microacontias_litoralis", "Acontias_litoralis", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Cryptoblepharus_boutonii", "Cryptoblepharus_metallicus", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Gallotia_gomerana", "Gallotia_bravoana", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Liolaemus_zullyi", "Liolaemus_zullyae", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Podarcis_atrata", "Podarcis_liolepis", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Laudakia_stellio", "Stellagama_stellio", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Bradypodion_caffrum", "Bradypodion_caffer", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Cordylus_polyzonus", "Karusasaurus_polyzonus", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Lygosoma_afrum", "Mochlus_sundevallii", lepidosaurtree$tip.label)
lepidosaurtree$tip.label <- gsub("Tupinambis_merianae", "Salvator_merianae", lepidosaurtree$tip.label)

# check$tree_not_data[check$tree_not_data == "Tapinurus_scutipunctatus"]
# ^ this code checks to see if a particular species name is in the phylogeny
```

## For species that are missing from our phylogeny, I have added them into our tree based on other phylogenies in the literature
## Using the bind.tip function, I've specied the name of the new species, the tip I want to add it to, and an arbitrary branch length
```{r}

lepidosaurtree <- bind.tip(lepidosaurtree, tip.label = "Bradypodion_kentanicum",
                           where = 2446, position = 0.1)
lepidosaurtree <- bind.tip(lepidosaurtree, tip.label = "Liolaemus_mapuche",
                           where = 2768, position = 0.1)
lepidosaurtree <- bind.tip(lepidosaurtree, tip.label = "Bradypodion_caeruleogula",
                           where = 6618, position = 0.5)
lepidosaurtree <- bind.tip(lepidosaurtree, tip.label = "Xenosaurus_newmanorum",
                           where = 3341, position = 0.1)
lepidosaurtree <- bind.tip(lepidosaurtree, tip.label = "Anolis_pentaprion",
                           where = 2970, position = 0.1)

# This code extracts the tip number for a species - which(lepidosaurtree$tip.label == "Aspidoscelis_flagellicauda")
# This code searches thre tree based on a sequence of nodes/tips - zoom(lepidosaurtree, 2970:2971 , subtree = TRUE)
# This code gets the parent node to a specific node - getParent(lepidosaurtree, node = 2768)
# This code searches the tree based on a character string such as a genus - zoom(lepidosaurtree, grep("Bradypodion", lepidosaurtree$tip.label))

```

## Now that we've added missing species and corrected species names, lets recheck mismatches between species in my data and phylogeny 
```{r}
check <- name.check(phy = lepidosaurtree, data = max_bf_overall, data.names = max_bf_overall$BinomialReptileDatabase)
```

## Drop species missing from the tree which aren't in our data using drop.tip
```{r}
lepidosaurtree <- drop.tip(lepidosaurtree, check$tree_not_data)
```

## Remove species from the data which are not in the tree
```{r}
matches <- match(max_bf_overall$BinomialReptileDatabase, check$data_not_tree, nomatch = 0)
max_bf_overall <- subset(max_bf_overall, matches == 0)
```

## Now, final recheck of the mismatches between species in my data and phylogeny - this should be "ok" if there are no mismatches! i.e. data and phylogeny contain the same species!
```{r}
check <- name.check(phy = lepidosaurtree, data = max_bf_overall, data.names = max_bf_overall$BinomialReptileDatabase)
```

# Last but not least, lets combine the phylogeny with our data!
# I've created 3 objects that combine the data with the phylogeny
```{r}
# This object contains all of the data
lepidosaur <- comparative.data(phy = lepidosaurtree, data = max_bf_overall, 
                               names.col =BinomialReptileDatabase, vcv = TRUE,
                                warn.dropped = TRUE, na.omit = FALSE)


# There is limited data for Sphenedontia (n = 1) and Gekkota (n = 2) for the superfamily analysis
# Subset the data to exclude Sphenodontia and Gekkota

max_bf_no_sphenodontia_gekkota <- max_bf_overall %>%
  filter(HigherTaxonomy != "Sphenodontia")%>%
  filter(HigherTaxonomy != "Gekkota") %>%
  droplevels()

# Create another object without Sphenodontia and Gekkota

lepidosaur_no_sphenodontia_gekkota <- comparative.data( phy = lepidosaurtree, data = max_bf_no_sphenodontia_gekkota, 
                               names.col =BinomialReptileDatabase, vcv = TRUE,
                                warn.dropped = TRUE, na.omit = FALSE)

# Check which species have been dropped
lepidosaur_no_sphenodontia_gekkota$dropped$tips
lepidosaur_no_sphenodontia_gekkota$dropped$unmatched.rows

# I need to remove Anguimorpha as there is limited body mass data for this superfamily
# Subset the data again to exclude Sphenodontia, Gekkota and Anguimorpha

max_bf_no_sphenodontia_gekkota_anguimorpha <- max_bf_overall %>%
  filter(HigherTaxonomy != "Sphenodontia")%>%
  filter(HigherTaxonomy != "Gekkota") %>%
  filter(HigherTaxonomy != "Anguimorpha") %>%
  droplevels()

# Create another comparative data without Sphenodon and Gekkota
lepidosaur_no_sphenodontia_gekkota_anguimorpha <- comparative.data(phy = lepidosaurtree, data = max_bf_no_sphenodontia_gekkota_anguimorpha, 
                               names.col =BinomialReptileDatabase, vcv = TRUE,
                                warn.dropped = TRUE, na.omit = FALSE)

# Double check to make sure the correct species have been dropped
lepidosaur_no_sphenodontia_gekkota_anguimorpha$dropped$tips
lepidosaur_no_sphenodontia_gekkota_anguimorpha$dropped$unmatched.rows


```
